---
output: 
  stevetemplates::syllabus:
    latex_engine: xelatex
    dev: cairo_pdf
geometry: margin=1in

title: "EH6105: Quantitative Methods"
author: Steven V. Miller
date: "Fall 2023"

# coteaching: TRUE

email: "steven.miller@ekohist.su.se"
officehours: "*on student request*"
office: "Södra Huset A 916"

classroom: "*varied* (S), E347 (D)"
classhours: "*check schedule*"
web: "eh6105.svmiller.com"

#fontfamily: mathpazo
fontfamily: cochineal
fontsize: 11pt
header-includes:
   - \linespread{1.05}
   - \LTcapwidth=.95\textwidth
---
 <!-- CCCCCCCCnncdxxxxxxx -->
<!-- Fun fact: my cat added this comment as she walked over the keyboard while I was working on this template. -->
<!-- I'm going to keep it going forward. Long live Queen Bea. -->
```{r setup, include=FALSE}
# knitr::opts_chunk$set(cache=FALSE, dev='pdf')
knitr::opts_chunk$set(cache=F,
                      fig.path = 'figs/',
                      cache.path='cache/',
                      warning=F,
                      message=F)

options(knitr.kable.NA = '')

knitr::opts_chunk$set(
                  fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      }
                  )


advdate <- function(obj, adv) {
 tmon <- obj + 7*(adv-1)
 tfri <- obj + 4 + 7*(adv-1)
 tmon <- format(tmon, format="%m/%d")
 tfri <- format(tfri, format="%m/%d")
 zadv <- sprintf("%02d", adv)
 tmp <- paste("Week ",zadv,sep='',", ", tmon," - ",tfri)
 return(tmp)
}

options(scipen=999)
#library(tidyverse)
#library(stevemisc)
#library(ggpmisc)
#library(anonymizer)
# library(ggcal)
#library(stringr)
#library(kfigr)
#library(broom)
#library(lubridate)



# library(RefManageR)
# library(knitcitations)
# library(rcrossref)
#bib <- ReadBib("~/Dropbox/master.bib")
#myopts <- BibOptions(bib.style = "authoryear", style="latex", first.inits=FALSE, max.names = 20)

# Create a calendar for your syllabus ----
# Source: http://svmiller.com/blog/2020/08/a-ggplot-calendar-for-your-semester/

# 1) what is the first Monday of the semester?
# Any number of ways to identify dates in R, but we'll use {lubridate} and the ymd() function here.
# Format: YYYYMMDD. In this example, 4 January 2021.

library(stevemisc)
library(bib2df)
library(tidyverse)
library(kableExtra)
library(stevethemes)

bib <- bib2df("~/Dropbox/master.bib")


mon <- ymd(20231127)
semester_dates <- seq(ymd(20231101), ymd(20240131), by=1)

seminar_dates <- c(
  ymd(20231130),
  ymd(20231201),
  ymd(20231205),
  ymd(20231208),
  # ymd(20231210),
  ymd(20231213),
  ymd(20231215),
  ymd(20231221),
  ymd(20240108),
  ymd(20240112)
)

lab_dates <- c(
  ymd(20231207), #x2
  ymd(20231211), #x2
  ymd(20231222), #x2
  ymd(20240103), #x2
  ymd(20240104), #x2
  ymd(20240110)  #x2
)

final_thing_due <- ymd(20240112)

problem_set_dates <- c(
  ymd(20231209),
  ymd(20231212),
  ymd(20231223),
  ymd(20240105),
  ymd(20240111)
)


# seminar_dates <- c(ymd(20230929), ymd(20231009), ymd(20231018))
# lab_dates <- c(ymd(20231003), ymd(20231006),
#                ymd(20231016), ymd(20231023),
#                ymd(20231027))
# problem_set_dates <- lab_dates + 1
# problem_set_dates[1:4] -> problem_set_dates
# rs_dates <- c(ymd(20231010), ymd(20231021))

tibble(date = semester_dates) %>%
  mutate(mon = lubridate::month(date, label=T, abbr=F), # get month label
         mon = fct_relevel(mon, "November", "December", "January"),
         wkdy = weekdays(date, abbreviate=T), # get weekday label
         wkdy = fct_relevel(wkdy, "Mon", "Tue", "Wed", "Thu","Fri","Sat"), # make sure Sunday comes first
         #week = wom(date),
         week = isoweek(date) - week(floor_date(date, unit = "months",
                                     week_start = 1)) + 1,
         semester = ifelse(date %in% semester_dates, 1, 0),
         lab = ifelse(date %in% lab_dates, 1, 0),
         day = lubridate::mday(date), # get day of month to add later as a label,
         seminar = ifelse(date %in% seminar_dates, 1, 0),
         category = case_when(
           #date %in% semester_dates ~ "Semester",
           date %in% seminar_dates ~ "Seminar",
           date %in% lab_dates ~ "Lab",
           #date %in% rs_dates ~ "Reading Summary Due",
           date %in% problem_set_dates ~ "Problem Set Due",
           TRUE ~ "NA",
         ),
         category = ifelse(date == final_thing_due, "Final Exam Due", category)) %>%
  ggplot(.,aes(wkdy, week)) +
  # custom theme stuff below
  # theme_steve_web() +
  theme_steve(style='ms') +
  theme(panel.grid.major.x = element_blank()) +
  # geom_tile and facet_wrap will do all the heavy lifting
  geom_tile(alpha=0.8, aes(fill=category), color="black", linewidth=.45) +
  facet_wrap(~mon, scales="free", ncol=3) +
  # fill in tiles to make it look more "calendary" (sic)
  geom_text(aes(label=day),family="Crimson Text") +
  # put your y-axis down, flip it, and reverse it
  scale_y_reverse(breaks=NULL) +
  scale_fill_manual(values=c("Seminar"="#dfc27d",
                             "Lab"="#bf812d",
                             #"Seminar/Lab"="#8c510a",
                             "Problem Set Due" = "#80cdc1",
                             "Final Exam Due" = "#415357",
                             "Reading Summary Due" = "#01665e",
                             "NA" = "white" # I like these whited out...
                             ),
                    breaks=c("Seminar","Lab", "Seminar/Lab", "Problem Set Due", "Reading Summary Due", "Final Exam Due")) +
  labs(x = "", y = "",fill="",
       caption = "All problem sets (and the final exam) are due *before* 5 p.m. on that corresponding day. Please check TimeEdit as well for confirmation of the schedule summarized here.") -> class_cal

```

# Course Description

The aim of the course is to provide the students with basic knowledge about quantitative methods in the social sciences, though the topical focus will privilege applications in economic history and international relations. The course will cover both descriptive statistics (i.e. summarizing and visualizing information) and inferential statistics (i.e. making statements about a population based on a sample of it). The major methodological focus will be ordinary least squares (OLS) regression, the basic workhorse of applied quantitative methods, though the course will take care to discuss matters of quantitative research design more broadly. Upon completion of the course, students should be able to read, interpret and assess quantitative research at a fundamental level. The class will serve as foundation training for a student's ability to design and perform a quantitative study, e.g., write a master thesis using quantitative methods.

# Learning Outcomes

1. Read and critically evaluate research in economic history and international relations where quantitative methods are used.

1. Formulate statistical models and test hypotheses

2. Solve basic problems by reference to probability theory, correlation analysis, and regression.

3. Use statistical software and present results from analyses of source material.

4. Demonstrate an awareness of ethical aspects of research.

# Course Policies

This section of the syllabus details multiple policies that will be implemented in this class through the semester.

## Required Readings

### Books

```{r, echo = FALSE, results="asis", eval=T} 
bib %>% filter_refs(c("kellstedtwhitten2013fpsr", "gujarati2014ee")) %>% capture.output() %>% print_refs()
``` 

Both books are available for purchase at Akademibokhandeln and can also be borrowed from the library.

### Articles

Be advised that additional articles may be assigned for individual lectures. Where applicable, this information will be communicated under Planeringar in Athena. The student is responsible for regularly checking Athena for updates. All articles should be available from [the university library](https://www.su.se/stockholm-university-library/). Students who are unfamiliar with how to use the library, but know how to Google things, may find the following link (on my blog) to be useful.

- [Set Up Google Scholar to Find Class Readings on Your Syllabus](http://svmiller.com/blog/2017/07/set-up-google-scholar-to-find-class-readings/)

### Other Resources

Quantitative methods is a unique subset of social science pedagogy where you could truly teach yourself if you had a working internet connection. Unlike other topics/fields in the discipline, social scientists who teach students about research design and methodology are apt to share their notes, lecture slides, and discuss stuff on public social media. Coding-related inquiries are even more abundant on a place like [Stack Overflow](https://stackoverflow.com/). You could truly "just Google" your way through a lot of this material if you were diligent and receptive to consuming as much information as you can find.

The instructor for the labs is not bashful about using past guides and material he has written for public consumption and encourages students to engage with these resources. Please consult the material on his blog ([`svmiller.com/blog`](http://svmiller.com/blog)) and the material/code for previous undergraduate and graduate-level methods classes he has taught.[^links] Lab sessions will almost assuredly borrow, quite liberally, from these past courses and blog posts.

[^links]: See [`posc3410.svmiller.com`](http://posc3410.svmiller.com) and [`post8000.svmiller.com`](http://post8000.svmiller.com) for more.

## Attendance Policy

The course consists of nine seminars, and five lab sessions. Participation in both is mandatory and the policy regarding absences in this course is quite strict. Students may miss no more than one seminar or lab, no matter the circumstances. A student who misses a seminar or lab must complete a make-up assignment, the contents of which are at the discretion of the instructor and will necessarily vary by the nature of the particular seminar or lab that the student missed. This make-up assignment must be submitted *and* accepted for credit on Athena no later than when the final exam is due. In the event the original submission is not accepted for credit, the student must do the assignment again and fix whatever issues the instructor communicated to the student to ultimately achieving a passing grade from the instructor. The student is able to do this as many as times as necessary in order to get a passing grade on the make-up assignment. However, a failure to get a passing grade on the make-up assignment will prevent the student from getting a grade on the course. 

<!-- The final grade will not be registered until the make-up assignment is approved. -->

Absence from two or more seminars and/or labs, no matter the circumstances, means the student must retake the course. Students bear sole responsibility for self-reporting an absence to the instructor as the instructor will only record attendance for seminars and labs.

## Computer Labs

This course consists of nine seminars and five computer labs. <!-- The three seminars, for which attendance is assumed but not formally required, are scattered across different dates, times, and rooms. --> The computer labs, for which attendance is *mandatory*, are all in Södra Huset E 347. Each lab day will have two groups: Group A and Group B. Students may select into either group, though the lab instructor's preference is to treat Group A as a beginner's group and Group B as a more advanced group. The latter group will ideally proceed through a lab session at a quicker pace. Students will ideally self-select and avoid cramming into the first session, which is always at either nine or ten in the morning. The later sessions have the benefit of permitting a more relaxed morning for the student. Please don't make me assign you into groups by fiat and please respect the fire code for the room.

### Software for Computer Labs

The lab sessions themselves will take place in rooms with computers in them for the student to use, though there is a (reasonable, implicit) assumption that the student has a personal computer. Tablets are not advised for these purposes as it is difficult to install the required third-party software needed for this course, which would be a major issue no matter the third-party software the student used for statistical analysis. Tablets will also typically lack the kind of memory and processing power for computational uses like this.

Lab sessions and problem sets (more in the next section) will all be done in the `R` programming language. Students should download this free software programming language at [cran.r-project.org](http://cran.r-project.org) and install it on their personal computer. Binaries are available for Windows and Mac (even Linux, if that is the weapon of choice for the student).

- The `R` scripts I provide are designed to work on the student’s computer with minimal maintenance. This should be clear in each particular script.

- The lab instructor *strongly* encourage students to contact him to learn about the language. He will assume that not discussing `R` with him means the student is fluent with the software and capable of proceeding through the course without additional oversight.

- Though not strictly mandatory for the course, consider getting a graphical user interface (GUI) front-end for `R` to learn it better. The instructor recommends (and will use) RStudio, which is available for free at [`posit.co`](https://posit.co/download/rstudio-desktop/). Do note there is a paid option of Rstudio that you *do **not** want.* The paid version is for servers. You want the basic open source integrated development environment (IDE). *This is **free***. Do not pay for anything related to R or Rstudio since you do not need whatever product is available for purchase.

The lab instructor published a [beginner's guide to using `R`](http://svmiller.com/blog/2014/08/a-beginners-guide-to-using-r/) in 2014 when he first started to teach courses that forced students to use the `R` programming language. He has since streamlined the `R` requirements for this class, making that guide somewhat dated. You will need to install the following packages, which are illustrated here with the `R` commands to install them. 

```{r usethese, echo=TRUE, eval=FALSE}
install.packages("tidyverse")      # for most things workflow
# ^ This is a huge installation. It should take a while.
install.packages("peacesciencer")  # some peace science data
install.packages("stevedata")      # for toy data sets to use in-class
install.packages("stevemisc")      # for some helper functions
install.packages("stevetemplates") # for preparing reports
install.packages("lmtest")         # for model diagnostics
```

The aforementioned `R` packages are not exhaustive of all packages a student may use in this course, and the lab instructor reserves the right to ask students to install additional packages along the way. Optional packages will be used in the labs, but are not strictly mandatory.

The `{tidyverse}` package will easily be the most time-consuming package to install, and the one most likely to give some students a potential problem during the course of its installation. In the strictest sense of the word "mandatory", this package is not "mandatory". It is possible to achieve the same results of this package by using either other packages in R or some functions that are default in R. However, not using this singular package---itself a wrapper for package for dozens of other R packages---would require the student to either download and install other R packages or require the student to learn code that is *much* less legible and intuitive than `{tidyverse}` code. Downloading and installing `{tidyverse}` is ultimately worth the effort, especially for beginners.

Based on experience, students may expect the following issues if the installation of this package results in an installation error ("non-zero exit status"), contingent on their operating system.

- *Mac*: You probably need to update Xcode. Xcode is a developer tool suite for Apple, and many of the `{tidyverse}` packages require access to developmental libraries that, to the best of my understanding, are available in Xcode. In all likelihood, you’re a first-time user who has not had to think about software development (and so you haven’t updated Xcode since you first got your Macbook). You might have to do that here. 
- *Windows*: The Windows corollary to Xcode is Rtools, which you don’t have installed by default (because it’s not a Microsoft program, per se). You’ll need to install it. First, take inventory of what version of R you have (for the university’s computer labs, it should be 4.0.5). Go to this website (https://cran.r-project.org/bin/windows/Rtools/) and download the version of Rtools that corresponds with the version of R you have. Just click through all the default options so that it can install.
- *Linux*: If you self-select into being a Linux user, you are aware of the consequences of your actions and that you may need to install various developmental libraries by yourself. The saving grace is that R packages that fail to install for the absence of developmental libraries they need will explicitly tell you what libraries you need, which are (practically) always in your package manager for you to install. This will very much depend on what particular distribution of Linux you're using. If your distribution of Linux is a fork of Debian, the lab instructor has this guide for you based on his trials and tribulations over the years.[^myguide]
- *Chromebook*: Students who self-select into Linux know what they signed up for, but students who have Chromebooks do not know they signed up for Linux. Read this guide to further assist you.[^guidechrome] It is understandable that students purchase Chromebooks for university education because they anticipate only needing something with a keyboard that is capable of checking email and writing documents. It is unfortunate that Chromebooks are lousy computers for situations like this. Students may want to [invest in a free posit Cloud account](https://posit.cloud/) if this is the situation in which they find themselves. Apparently, the language studio has R and RStudio installed on their computers and those computer labs are more accessible than the rooms in which we'll meet (because they are open and E 347 is routinely closed).

[^myguide]: http://svmiller.com/blog/2019/07/notes-to-self-new-linux-installation-r-ubuntu/
[^guidechrome]: https://levente.littvay.hu/chromebook/



## Assessment

Course assessment consists of five problems sets in the `R` programming language along with a written final exam. The problem sets revolve around some `R` scripts I will provide. The questions in these scripts will tell students to answer questions based on the output of a function in the script, or prompt students to extend the script in some manner (i.e. to induce some programming skills from the student). Problem sets must be typed in either a Word processor or, preferably, La\TeX. Each problem set is graded on a 0-8 scale, resulting in a maximum of 40 points students will earn toward their grade. There is an additional final exam, worth 60 points, that will be given at the end of the semester. This results in 100 total points for the semester.

Computer lab sessions are mandatory and students must attend one of the two sessions that day to receive credit for attendance. The instructor will take attendance by making students sign their name on attendance sheet, making it the student's responsibility to ensure their signature is next to their name by the end of the lab session in order to receive credit. Failure to attend a lab session---even under extraordinary circumstances---will require a complementary assignment to be completed *in addition to* the problem set associated with the lab session. The nature of this assignment will be at the instructor's discretion. Do note that missing more than one lab, under any circumstance, will result in the student's inability to finish the course.

Do note the nature of the problem sets and the exam preclude students from sharing answers. Such a form of cooperation will be considered plagiarism. All problem sets and the exam should be submitted through Athena *before* the 5 p.m. deadline each day. Resist the urge to assume your laptop's time is the system's time, and that submission at the literal last minute is an acceptable strategy.

```{r, echo=F, eval=T, results="asis"}

tribble(~a, ~b, ~c,
        "Assignments (5x)", 5, 40,
        # "Randomly Selected Assignment", 10, 10,
        "Final Exam/Paper", 60, 60,
        "TOTAL", NA, 100
        )  %>%
  kbl(.,
      align = "lcc",
      longtable = T, booktabs = T, linesep = '',
      caption = "A Summary of the Assessment in This Course",
      col.names = c("Item", "Points (Each)", "Total Points")) %>%
  row_spec(0, bold=TRUE) %>%
  row_spec(2, hline_after = TRUE) %>%
  row_spec(3, italic = TRUE, background = "gray!10")
```

## Plagiarism

Students must be conscientious about how they use the material they read and how they summarize them for in-class assignments. Simply put, the use of other people's materials without attribution is plagiarism. The translation of someone else's materials into another language, without attribution, is also plagiarism. A verbatim copy of someone else's words, with selective changes of just a few words, is also plagiarism when done without attribution. This form of academic dishonesty is considered cheating and exposes the student to a failing grade and potential disciplinary measures on top of that. Any student who is caught cheating or disrupting academic activities can be suspended from lectures and exams for a period of up to six months. The Vice-Chancellor or the Disciplinary Council decides whether the student is to be subject to any disciplinary measures. Please consult [the Student Support portal at Stockholm University](https://www.su.se/english/education/student-support/your-rights-and-responsibilities) for more information about your rights, and your responsibilities as students.

## Language

The instructor does not yet speak Swedish, but is working on it. The seminars, labs, and all course material will be in English. All assignments must be completed in English as well.

## Course Evaluation

Students can evaluate the instructor and the course through a written evaluation, which will be distributed at the end of the course. The course evaluation itself is voluntary, but strongly encouraged as this important form of feedback shapes course development in the department. Student responses are 100% anonymous.



### Grading Criteria

Table 2 provides a summary of the distribution of possible grades for the semester, along with a description of what each grade suggests in terms of the learning outcomes.

```{r, echo=F, results="asis"}

tribble(~grade, ~points, ~summary, ~description,
        "F", "0-39", "FAIL", "The student does not reach the requirements for E.",
        "Fx*", "40-49", "INSUFFICIENT", "The student does not reach the requirements for E, but get the opportunity to provide supplementary work to reach E.",
        "E", "50-59", "ADEQUATE", "The student demonstrates a limited theoretical knowledge and ability to apply some of the expected learning outcomes. The student can formulate and test some basic statistical hypotheses, solve problems using probability theory, correlation, and regression analysis. The student also demonstrates knowledge of ethical aspects in quantitative research and show a limited command of R to present quantitative data descriptions and analyses.",
        "D", "60-69", "SATISFACTORY", "The student demonstrates a satisfactory theoretical knowledge and limited ability to apply several expected learning outcomes. The student can formulate and test basic statistical hypotheses, solve problems using probability theory, correlation, and regression analysis. The student also demonstrates knowledge of ethical aspects in quantitative research and show a limited command of R to present quantitative data descriptions and analyses",
        "C", "70-79", "GOOD", "The student demonstrates a good theoretical knowledge and ability to apply a majority expected learning outcomes. The student can formulate and test statistical hypotheses, solve problems using probability theory, correlation, and regression analysis. The student also demonstrates awareness of ethical aspects in quantitative research and can use R to present quantitative data descriptions and analyses.",
        "B", "80-89", "VERY GOOD", "The student demonstrates a very good theoretical knowledge and ability to apply all expected learning outcomes. The student can, in an independent manner formulate and test statistical hypotheses, solve problems using probability theory, correlation and regression analysis. The student also demonstrates good awareness of ethical aspects in quantitative research and shows good command of R to present quantitative data descriptions and analyses.",
        "A", "90-100", "EXCELLENT", "The student demonstrates an exceptionally good theoretical knowledge and ability to apply all expected learning outcomes. The student can, in a critical and independent manner formulate and test statistical hypotheses, solve problems using probability theory, correlation and regression analysis. The student also demonstrates and excellent awareness of ethical aspects in quantitative research and shows very good command of R to present quantitative data descriptions and analyses.") %>%
  kbl(.,
      align = "cccl",
      longtable = T, booktabs = T, linesep = '',
      caption = "A Description of the Grading Outcomes for the Semester",
      col.names = c("Grade", "Point Range", "Summary", "Description")) %>%
  kable_styling(latex_options = "striped", stripe_color = "gray!20") %>%
  column_spec(4, width = "24em") %>%
  footnote(symbol = "A complementary assignment given by the examiner must be handed in within a week to get the grade E.") %>%
  kable_styling(latex_options = c("repeat_header")) %>%
  row_spec(0, bold=TRUE)# %>%
#  row_spec(c(1,3,5,7), hline_after = TRUE) %>%
 # row_spec(c(2, 4, 6, 8), hline_after = TRUE)

```


\newpage

# Class Schedule

The following serves as the schedule for this course, structured by seminar dates and topics and computer lab dates and topics. Seminar locations are varied but *all computer labs will be in Södra Huset E 347*. Each section heading communicates the topic and, in the case of the seminars, what readings from the Kellstedt and Whitten (2013) [`KW`] and Gujarati (2014) [`G`] should be done *before* the seminar. Any additional article readings will be posted to Athena in due course, when applicable. Information about the time and location of the seminars follow the section heading and description of the seminar. Students should have this information in TimeEdit as well. In the event of a miscommunication about the time and location of a seminar, the information on TimeEdit should be considered correct and the instructor will issue a clarification on Athena.


<!-- Readings from the Gujarati (`G`) text are also listed in the section heading and should be read *before* the seminar.  Students should have this information in TimeEdit as well. In the event of a miscommunication about the time and location of a seminar, the information on TimeEdit should be considered correct and the instructor will issue a clarification on Athena. -->


<!-- Each section heading communicates the topic, time and date of the meeting, and the instructor overseeing the meeting.  Karina Shyrokykh (`KS`) will oversee the seminars and Steve Miller (`SM`) will oversee the labs. Readings from the Warne (`W`) text are also listed in the section heading and should be read *before* the seminar. Additional article readings will be posted to Athena in due course, when applicable. *All seminars will be in Södra Huset B 705 and all computer labs will be in Södra Huset E 347*. Information about the time and location of the seminars follow the section heading and description of the seminar. Students should have this information in TimeEdit as well. In the event of a miscommunication about the time and location of a seminar, the information on TimeEdit should be considered correct and the instructor will issue a clarification on Athena. -->

```{r calendar, echo=F, eval=T, fig.cap="A Calendar for the Course", results="hide",  warning=F, fig.width=8, fig.height=4.5}
class_cal
```



## Seminar 1, 30/11: Introduction (KW, chp. 1)

*This will be an introductory lecture introducing students to the course and its requirements.*

**Time and location**: Södra Huset F 389, 15-17

## Seminar 2, 01/12: Research Design (KW, chps. 2-4)

*This lecture will discuss basic elements of a research design, including the conceptualization of topics and operationalization of concepts into variables.*

**Time and location**: Södra Huset F 289, 10-12

## Seminar 3, 05/12: Descriptive Statistics (KW, chp. 6)

*This lecture will introduce students to some basic descriptive statistics, focusing on quantitative assessments of central tendencies in data and the variation in variables.*

**Time and location**: Södra Huset B 413, 15-17

## Computer Lab 1, 07/12: Introduction to R

*This lab will introduce students to the R programming language and Rstudio. This will be more of an orientation than anything, though students will ideally also learn about reading data as well as some basic data manipulation.*

- Group A: 9-11
- Group B: 11-13

## Seminar 4, 08/12: Probability and Sampling (Chp. 7)

*This is a lecture of two halves. In the first part, students will get a crash course on what is probability and some rudimentary aspects of probability theory. In the second part, students will learn about the logic of random sampling from a population in order to make inferential statements about the population in question.*

**Time and location**: Södra Huset D 299, 10-12

## Computer Lab 2, 11/12: Descriptive Statistics

*This lab will go over some basics of descriptive statistics. Topics will be multiple, though the basic takeaway will implore students to first look at their data before analyzing it. Descriptive statistics is also good diagnostics.* 

- Group A: 10-12
- Group B: 13-15

## Seminar 5, 13/12: Hypothesis Testing (KW, Chp. 8)

*This lecture will dovetail the previous lecture on probability and sampling to its application in evaluating hypotheses.*

**Time and location**: Södra Huset B 413, 15-17

## Seminar 6, 15/12: Bivariate OLS (KW, Chp. 9)

*Here, we introduce students to the basics of ordinary least squares (OLS) regression in the case of two variables. This may seem daunting to students to start, but students who remember the slope-intercept equation from their secondary education should see the basic intuition here.*

**Time and location**: Södra Huset B 497, 10-12


## Seminar 7, 21/12: Extending OLS Regression (KW, Chp. 10; G, chps. 1-3)

*Once students get the basics of OLS in the bivariate case, extending it becomes a somewhat trivial matter. Topics here will include multiple regression, interactions, fixed effects, and more.*

**Time and location**: Södra Huset D 299, 10-12


## Computer Lab 3, 22/12: OLS Regression

*This lab will go over basic OLS regression, focusing on the simple bivariate case.*

- Group A: 9-11
- Group B: 11-13

## Computer Lab 4, (03/01; 01/04): OLS Regression 2

*This lab will extend OLS regression to multiple regression, including so-called control variables, fixed effects, and more. What emerges is still the familiar "rise over run" of the slope-intercept equation, at least as students should interpret it.*

- Group A: 10-12 (03/01)
- Group B: 10-12 (04/01)

## Seminar 8, 08/01: Testing Assumptions (KW, chp. 11; G, chps. 4, 5, 7)

*All models are valid if and only if the assumptions of the model hold. This lecture will expand on the basic assumptions of OLS and identify cases where OLS is a questionable or even invalid model for a given application.*

**Time and location**: Södra Huset F 389, 15-17

## Computer Lab 5, 10/01: Testing Assumptions

*This lab will over the basic diagnostic tests of the OLS model, and what you should do if your given model fails one of those diagnostic tests.*

- Group A: 10-12
- Group B: 13-15

## Seminar 9, 12/01: Causality, Its Issues, and Other Matters (KW, chps. 3-4)

*Our final lecture will talk about the deceptive difficulty that comes in trying to assess neat, causal relationships in a world governed by endogeneity. TIme and space permitting, we can use this as a sort of Q&A for students to ask questions about the material.*

**Time and location**: Södra Huset C 497, 10-12